---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
      
### Loading and preprocessing the data 
```{r, echo=TRUE}
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date)
summary(activity)
attach(activity)
```



### What is mean total number of steps taken per day? 


1. Calculate the total number of steps taken per day
```{r, echo=TRUE}
steps.perday <- sapply(split(activity, date), function(a) sum(a$steps,na.rm=T) )
steps.perday
```

2. Make a histogram of the total number of steps taken each day
```{r, echo=TRUE}
hist(steps.perday, xlab="steps", breaks=10,
     main="Histogram of the total number of steps per day")
```

3. Calculate and report the mean and median of the total number of steps taken per day
```{r, echo=TRUE}
mean(steps.perday)
median(steps.perday)
```



### What is the average daily activity pattern?

1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r, echo=TRUE}
daily.activity <- sapply(split(activity, interval), function(a) mean(a$steps,na.rm=T) )
plot(names(daily.activity), daily.activity, type="l", 
     xlab="clock", main="Average Daily Activity Pattern")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo=TRUE}
names(sort(-daily.activity))[1]
```
The 5-minute interval which contains the maximum number of average steps is 8:35.



### Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r, echo=TRUE}
good <- complete.cases(activity)
nrow(activity)-nrow( activity[good,] )
sum(is.na(steps))    # Double Check
```
The total number of missing values in the dataset is 2304.

2. Devise a strategy for filling in all of the missing values in the dataset. Here I use the median for that interval.
```{r, echo=TRUE}
imp <- sapply(split(activity, interval), function(a) median(a$steps,na.rm=T) )
imp
```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r, echo=TRUE}
activity.imp  <- activity
detach()
attach(activity.imp)
for (i in 1:nrow(activity.imp) )  {
      steps[is.na(steps)] <- imp[ as.character( interval[is.na(steps)] ) ]
}
activity.imp$steps <- steps
summary(activity.imp)
```
activity.imp is the new dataset with the missing data filled in.


4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r, echo=TRUE}
steps.perday.imp <- sapply(split(activity.imp, date), function(a) sum(a$steps,na.rm=T) )
mean(steps.perday.imp)
median(steps.perday.imp)
summary(steps.perday)
summary(steps.perday.imp)
steps.perday.imp[!steps.perday==steps.perday.imp]
hist(steps.perday.imp, xlab="steps", breaks=10,
     main="Histogram of the total number of steps per day(after imputing NAs)" )
```     

The mean of the total number of steps per day is slightly greater than the orginal one, while the median remains the same. The histogram of total number of steps per day remains the same as the original one. Imputing missing data have only a tiny impact on the estimates of the total daily number of steps, since only 8 days of NA values are imputed, and imputed values are very small, only influencing the lower tail of the distribution.



### Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r, echo=TRUE}
wkd <- weekdays(date)
for (i in 1:length(wkd) )  {
      if (wkd[i]=="Sunday" | wkd[i]=="Saturday")   wkd[i] <- "weekend"
      else   wkd[i] <- "weekday"
}
activity.imp$wkd <- as.factor(wkd)
summary(activity.imp)
```
Variable wkd indicates whether a given date is a weekday or weekend day.


2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r, echo=TRUE}
activity.imp.wkday <- subset(activity.imp, wkd=="weekday")
activity.imp.wkend <- subset(activity.imp, wkd=="weekend")
str(activity.imp.wkday)
str(activity.imp.wkend)
daily.activity.wkday <- sapply(split(activity.imp.wkday, interval), function(a) mean(a$steps) )
daily.activity.wkend <- sapply(split(activity.imp.wkend, interval), function(a) mean(a$steps) )
par(mfrow = c(2, 1))
plot(names(daily.activity.wkday), daily.activity.wkday, type="l", ylim=c(0,200),
     xlab="clock", main="Average Daily Activity Pattern of weekdays")
plot(names(daily.activity.wkend), daily.activity.wkend, type="l", ylim=c(0,200),
     xlab="clock", main="Average Daily Activity Pattern of weekends")     
```

Above is the panel plot made by Base Plotting System.


```{r,echo=TRUE}
daily.activity.imp <- c(daily.activity.wkday,daily.activity.wkend)
daily.activity.interval <- rep(unique(activity$interval),2)
daily.activity.wkd  <- rep(c("weekday", "weekend" ), each=length(daily.activity.wkday))
library(lattice)
xyplot(daily.activity.imp~daily.activity.interval |daily.activity.wkd , type="l", 
       layout = c(1, 2), xlab="intervals", ylab="number of steps")
```

Above is the panel plot made by Lattice Plotting System.


